---
title: An Analysis of the Properties of the Exponential Distribution and Demonstration
  of the CLT
author: "Connor M"
date: "March 14, 2015"
output:
  pdf_document:
    keep_tex: yes
---

## Introduction  

It will be the purpose of this report to compare and contrast the following:  

1. The mean of a sample drawn from an exponential distribution and the mean of the theoretical exponential distribution said sample was drawn from.  
2. The variance of a sample drawn from an exponential distribution and the variance of the theoretical exponential distribution said sample was drawn from.  

We will also show how the distribution of a large amount of sample means drawn from an exponential distribution approximately follows a gaussian (normal) distribution. This phenomenon is described by the Central Limit Theorem (CTL).  

## Sample Mean vs. Population Mean
*Note: For the purposes of this report, the distribution for which all samples will be drawn from will be an exponential distribution with $\lambda = 0.2$. i.e. $X \sim Exp(\lambda = 0.2)$*

The expected value (or mean) of an exponential distribution is $E(x) = \frac{1}{\lambda}$. Since $\lambda = 0.2$ for our theoretical distribution, the mean will be 5. Similarly, the standard deviation of our theoretical distribution is also $\frac{1}{\lambda}$ therefore so $\sigma = 5$.  
  
\[\mu = E(X) = \frac{1}{\lambda} = \frac{1}{0.2} = 5\]  

\[\sigma = \frac{1}{\lambda} = \frac{1}{0.2} = 5\]  

Our first task is to compare the theoretical (or population) mean of our exponential distribution to the mean of a sample drawn from our population distribution.

Let's first start with 

```{r initialize}
set.seed(1234)
```

```{r init_sample}
sample <- rexp(n = 1000, rate = 0.2)
```

```{r x_bar}
sample_mean <- mean(sample)
```

```{r sigma}
sample_sd <- sd(sample)^2
```

```{r}
sample_matrix <- matrix(rexp(40000, 0.2), 1000)

average_means <- data.frame(avg_means = rowMeans(sample_matrix))
```

```{r}
library(ggplot2)

ggplot(average_means, aes(x = average_means$avg_means)) + 
    geom_histogram(colour = "black", fill = "white") +
    ylab('Frequency') +
    xlab('Sample Mean Values')
```

```{r}
standardized_means = (average_means - 5) / 5

ggplot(standardized_means , aes(x = standardized_means$avg_means))  + 
    geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
    stat_function(fun=dnorm, aes(x = rnorm(n = 1000, mean = 5, sd = 5)), color = "#FF6666") +
    ylab("Density") +
    xlab("Sample Mean Values")
```
